---
title: "Project MUTI Semester 1:MSc (Data Handling Submission Using R)"
author: "Gagan Deep Singh Chhabra"
date: "December 8, 2016"
output: html_document
---

# Remove any previous loaded objects in the library 
--------------------------------------------------
```{r}
rm(list=ls())
ls()
```
# Setting up working directory
```{r}
setwd("D:\\R\\MUIT\\stats2-master\\MUIT_STAT_PROJECT2")
```
## Created a student data with 50 Records.
# First	Import the Data Set and display first few rows convert all missing values to NA ---------------------------------------------------------------------------------------
```{r}
student1 = read.csv('StudentProjectData.csv',stringsAsFactors = F,na.strings=c("",".","NA"))
head(student1)
```

## To see the data in Table form
```{r}
edit(student1)
```

# 2	Check and report structure, Dimensions, Column Names of DF --------------------------------------------------------------
```{r}
str(student1)
dim(student1)
names(student1)
colnames(student1)
length(student1)
nrow(student1)
```
# Now check the if Name of Student have missing values and than replace with somename

```{r}
student1$name = NULL
str(student1)
length(student1)
```

# Create a vector and insert new names of Students 
--------------------------------------------------
```{r}
stuname = paste('S-',seq(1:nrow(student1)),sep='')
stuname
student2 = cbind(student1, I(stuname))
str(student2)
student2$stuname
```
# Saving file for records
```{r}
write.csv(student2,"StudentProjectDatabackup.csv")
```
# Assign sname  nos as rownames and delete stuname --------------------------------------------------
```{r}
rownames(student2) = student2$stuname
student2$stuname = NULL
str(student2)
rownames(student2)
```


# List Names of Students having missing values 
----------------------------------------------
```{r}
student2[!complete.cases(student2),] # this will list only where NA is there
```
# There are missing values also - should have converted them into NA at time of import

```{r}
student2[is.na(student2)]
```


# List rows having missing values and count them -----------------------------------------
```{r}

student2[!complete.cases(student2),]
sum(is.na(student2)) # no of rows having NA Values
```

## We found 30 rows with null value

# List Rows without Missing Values (First 5 colns) & total them ---------------------------------------------------------------
```{r}
student2[complete.cases(student2),1:5]
sum(complete.cases(student2)) # No of rows having complete values
```


# List Columns having Missing values 
--------------------------------------
```{r}
colnames(student2)[colSums(is.na(student2)) > 0] # only relevant columns
```
# another way
```{r}
apply(is.na(student2),2,any) # True colns have NA values
colnames(student2)[apply(student2,2,anyNA)] # Only colns having NA 
```
# We found following columns with NA
# "feepaid" "Gender"  "Stats"   "Rprog"   "Vba"     "Sql"     "Grade"   

- First we need to analysis the data and than remove the NA with appropiate value or as per business decision


- Use a plot of fees paid to find out missing values
```{r}
plot(student2$feepaid)
```
# explain how
```{r}
names(student2)
```
# We found 4 outliers and two missing values NA
# we decided to replace NA with average value mean

# Missing values of fees with Mean of fees
```{r}
sum(is.na(student2$feepaid))
student2$feepaid[is.na(student2$feepaid)] = mean(student2$feepaid,na.rm=T)
```
# Gender Replacing NA with F - Female
```{r}
student2$Gender[is.na(student2$Gender)] <- "F"
```

# Stat, Rprog, Vba, Sql replacing with the average marks of students in the class
```{r}
student2$Stats[is.na(student2$Stats)]  = mean(student2$Stats[!is.na(student2$Stats)])
student2$Rprog[is.na(student2$Rprog)]  = mean(student2$Rprog[!is.na(student2$Rprog)])
student2$Vba[is.na(student2$Vba)]  = mean(student2$Vba[!is.na(student2$Vba)])
student2$Sql[is.na(student2$Sql)]  = mean(student2$Sql[!is.na(student2$Sql)])
```
# Replacing grade by A
```{r}
student2$Grade[is.na(student2$Grade)] = "A"
edit(student2)
```
# Double check the rows for NA Values
```{r}
student2[!complete.cases(student2),] # this will list only where NA is there
```
# Data is now clean, all NA values have been removed and now ready for analysis

# Distribution of Grade and Gender 
-----------------------------------
```{r}
table(student2$Gender)
table(student2$Grade)
ftable(student2$Gender,student2$Grade)  # cross table table
```
## Analysis Females are more intelligent and good in studies as compare to Male Below is the analysis
#     A  B
# F  20  0
# M  23  7


# Display first 5 and last 5 rows together 
------------------------------------------
```{r}
rbind(head(student2[,1:8]),tail(student2[,1:8]))
```
# Check to see if duplicate rollnos 
-----------------------------------
```{r}
duplicated(student2$rollno)
```
## No duplicate record found

# Edit data within R Change year of batch to 2014 
-------------------------------------------------
```{r}
student2 = edit(student2)
student2[1:2,]
```
# One Last Final Check of incomplete data
```{r}
student2$rollno[complete.cases(student2)]
```
# Using plyr::count count no of students gender - Class wise 
```{r}
library(plyr)
count(student2,c('Gender','Class'))
```
# Analysis
#       Gender Class freq
#   1      F   BSc    8
#   2      F   MSc   12
#   3      M   BSc   12
#   4      M   MSc   18

#list students have Sql Marks  > 90
```{r}
student2[student2$Sql > 90,1:12]
```


# ********************************************
# Average marks of all subjects of all students
```{r}
mean(student2$Stats)  # 83.91304
mean(student2$Rprog)  # 83.91304
mean(student2$Vba)    # 84.375
mean(student2$Sql)    # 83.91304


apply(student2[,c('Stats','Rprog','Vba','Sql')],2,mean,na.rm=T) # ColSums
apply(student2[,c('Stats','Rprog','Vba','Sql')],1,mean,na.rm=T) # RowSums
```


# Create another column average marks of all subjects
```{r}
student2$MarksAverage <- rowMeans(student2[,9:12])
```
# Sort, Rank Order of Students 
------------------------------
```{r}
rank(student2$MarksAverage,ties.method = 'average')
student2[50,c('RollNo','MarksAverage')]
sort(student2$Class,decreasing = T)
student2[with(student2,order(Class,MarksAverage)),c('RollNo','Name','Class','Gender','Batch')]

edit(student2)
```
# Save data, functions  in Rdata object 
---------------------------------------
```{r}
save(student2,file='gagan.Rdata')
```
# Save History in a file your firstname.History
-----------------------------------------------

savehistory(file='gagan.Rhistory')

